---
title: "R/Bioconductor tools for reproducible proteomics data analysis"
author: "Laurent Gatto"
date: "11/5/2018"
output: html_document
---

> Find this documet at [https://github.com/lgatto/2018_11_05_RBiocProt_DDUV](https://github.com/lgatto/2018_11_05_RBiocProt_DDUV)

# Introduction

1. Use case: Quantitative MS-based spatial proteomics (intro from
   [https://zenodo.org/record/1435058](https://zenodo.org/record/1435058)).
2. [Data structure](http://lgatto.github.io/pRoloc/articles/Figures/msnset.png).


# Demo

## Working environment

```{r env}
suppressPackageStartupMessages(library("MSnbase"))
library("pRoloc")
library("pRolocdata")
setStockcol(NULL)
library("pRolocGUI")
library("magrittr")
```

## Loading data into R

```{r in}
f0 <- dir(system.file("extdata", package = "pRolocdata"),
	  full.names = TRUE,
	  pattern = "Dunkley2006")
basename(f0)

res <- readMSnSet2(f0, ecol = 5:20)
res
```

## Accessing data and meta-data


```{r slots}
data(hyperLOPIT2015)
x <- hyperLOPIT2015
exprs(x)
pData(x)
fData(x)
```

## Missing data

```{r na}
data(naset)
naplot(naset)

filterNA(naset, pNA = 0.5) %>%
    naplot

naset %>%
    filterNA( pNA = 0.5) %>%
    impute(method = "knn") %>%
    naplot

res <- naset %>%
    filterNA( pNA = 0.5) %>%
    impute(method = "knn")
```

## Normalisation

```{r norm}
par(mfrow = c(1, 2))
boxplot(exprs(res))

res %>%
    normalise(method = "vsn") %>%
    exprs %>%
    boxplot
```

## Dimensionality reduction

```{r pca}
hl <- hyperLOPIT2015
plot2D(hl)


fvarLabels(hl)
plot2D(hl, fcol = "final.assignment")

ptz <- exp(fData(hl)$svm.score) - 1
plot2D(hl, fcol = "final.assignment", cex = ptz)

setStockcol(paste0(getStockcol(), 80))
plot2D(hl, fcol = "final.assignment", cex = ptz)
```

## Classification

```{r class}
data(tan2009r1)

## params <- svmOptimisation(tan2009r1, fcol = "markers.orig",
##               times = 100, xval = 5,
##               verbose = FALSE)

fn <- dir(system.file("extdata", package = "pRoloc"),
      full.names = TRUE, pattern = "params.rda")
load(fn, verbose = TRUE)
params

plot(params)
levelPlot(params)

svmres <- svmClassification(tan2009r1, fcol = "markers.orig",
			    sigma = 1, cost = 1)
fvarLabels(svmres)

svmres <- svmClassification(tan2009r1, fcol = "markers.orig",
			    assessRes = params)
processingData(svmres)
fvarLabels(svmres)
```

## Interactive visualisation


Offline

```{r gui, eval = FALSE}
pRolocVis(hl)
```

or online at [https://lgatto.shinyapps.io/christoforou2015/](https://lgatto.shinyapps.io/christoforou2015/).

# References

Breckels LM, Mulvey CM, Lilley KS and Gatto L. **A Bioconductor
workflow for processing and analysing spatial proteomics data**
[version 2; referees: 2 approved]. F1000Research 2018, 5:2926
(https://doi.org/10.12688/f1000research.10411.2)

`pRoloc` documentation: http://lgatto.github.io/pRoloc/

Gatto L, Lilley KS. MSnbase-an R/Bioconductor package for isobaric
tagged mass spectrometry data visualization, processing and
quantitation. Bioinformatics.  2012 Jan 15;28(2):288-9. doi:
10.1093/bioinformatics/btr645. Epub 2011 Nov 22.  PubMed PMID:
22113085.

`MSnbase` documentation: http://lgatto.github.io/MSnbase/

Gatto L, Breckels LM, Naake T, Gibb S. **Visualization of proteomics
data using R and bioconductor**. Proteomics. 2015
Apr;15(8):1375-89. doi: 10.1002/pmic.201400392. Review. PubMed PMID:
25690415; PubMed Central PMCID: PMC4510819.

Gatto L, Christoforou A. **Using R and Bioconductor for proteomics
data analysis**. Biochim Biophys Acta. 2014 Jan;1844(1 Pt
A):42-51. doi: 10.1016/j.bbapap.2013.04.032. Epub 2013
May 18. Review. PubMed PMID: 23692960.
